---
title: "Unsupervised classification"
author: "Andrea Huerfano"
date: "September 8, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r , echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
packages<-c("FactoMineR", "factoextra", "xtable", "NbClust", "plyr","GGally", "ggcorrplot")
lapply(packages, library, character.only=TRUE)
```
This dataset is a subset of the results of a chemical analysis of Italian wine. The analysis determined the quantities of 13 constituents of red wines. These are the variables:

+ fixed acidity
+ volatile acidity
+ citric acid
+ resigual sugar
+ chlorides
+ free.sulfur.dioxide
+ total.sulfur.dioxide
+ density
+ ph
+ sulphates
+ alcohol
+ quality

```{r,echo=TRUE, warning=FALSE, message=FALSE}
 winequality.red <- read.csv("C:/Users/Andrea/Desktop/wine/winequality-red.csv", sep=";")
```

Looking into the data structure we can see a strong direct/inverse relation between some variables like fixed-volatile acidity, free-total sulfor dioxide and alcohol-quality.

```{r, fig.height=3.5, fig.pos='h'}
corr <- round(cor(winequality.red), 1)
ggcorrplot(corr)
```

#Kmeans
##Number of cluster

The data is standardized before clustering to obtain a  better quality, efficient and accurate cluster result. The library *NbClust* is used to calculate the optimum number of clusters, we are going to use the Friedman index. 


```{r,echo=TRUE, warning=FALSE, message=FALSE}
df<-scale(winequality.red)
res.wine<-NbClust(df, distance = "canberra", min.nc = 2, max.nc = 10, method = "kmeans", index = "friedman")
res.wine$Best.nc
```

Another methods to select the number of cluster are plotted below.
```{r,echo=TRUE, warning=FALSE, message=FALSE,fig.height=2.5, fig.pos='h'}
set.seed(21)
elbow<-fviz_nbclust(df, kmeans, method = "wss") +geom_vline(xintercept = 4, linetype = 2)+ labs(subtitle = "Elbow method"); elbow

Silhouette<-fviz_nbclust(df, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method");Silhouette

gap<-fviz_nbclust(df, kmeans, nstart = 25,  method = "gap_stat", 
  nboot = 50)+labs(subtitle = "Gap statistic method"); gap
```

After all of the results, I decided to use the mode and in this way, obtaining that the better number of clusters is 4.

##Kmeans algorithm
No forget to set a seed to obtain every run the same results.
```{r, echo=TRUE}
set.seed(10)
n_cluster<-4
kmedia2<-kmeans(df, n_cluster, nstart = 1000)

```

Let see the centroids
```{r, echo=TRUE}
round(kmedia2$centers,2)
```
And the clusters over all of the variables are painted in the scatter plots when each color represents a different cluster.

```{r, fig.pos='h'}
plot(winequality.red, col=kmedia2$cluster, 
     main="Posición de los grupos respecto a las variables")
points(kmedia2$centers, col = kmedia2$cluster, pch = 23, cex = 1.5)
```

Too see in detail one of the scatterplots
```{r,fig.height=3, fig.pos='h'}
plot(winequality.red$chlorides, winequality.red$residual.sugar, col=kmedia2$cluster, xlab = "chlorides",ylab = "residual sugar")
legend(0,100, legend=c('1', '2', '3', '4'),col=c("black","blue", "red","green" ),
pch = 18, cex = 1)
```
It is really imporant look into the withinss and beteens sum of squares.
```{r}
print(round(kmedia2$withinss,2))
```

```{r}
print(kmedia2$betweenss)
```

In this case, the sum of squares over the betweens is greater than the within, which means that the segmentation creates cluster different between them but internally the observations are similar, homogeneity inside and heterogeneity outside the clusters. 

#Hierarchichal Clustering
##Principal Component Analysis
To use the function HCPC is necessary to run before a PCA, because the cluster are made over the space generated in the PCA.


```{r, fig.height=3.8, fig.pos='h'}
res.PCA<-PCA(df, ncp = 2)
```

It is necessary to determine the minimum and the maximum number of possible clusters. If nb.clust=-1 the optimal number of clusters is used and if nb.clust is an integer it fixes the number of clusters. The indexed hierarchy is represented by a tree named a dendrogramm.
```{r,fig.height=4, fig.pos='h'}
set.seed(10)
res.hcpc<-HCPC(res.PCA,  nb.clust = -1,min = 3, max = 6)
```


#Facto investigate
Finally, a quickly way to have an analysis of your results is using the function Investigate which generate an extra file with the analysis. 
```{r, warning=FALSE, message=FALSE}
library(FactoInvestigate)
Investigate(res.hcpc,document="pdf_document")
```


# References
 Sebastien Le, Julie Josse, Francois Husson (2008).
  FactoMineR: An R Package for Multivariate Analysis.
  Journal of Statistical Software, 25(1), 1-18.
  10.18637/jss.v025.i01

  Alboukadel Kassambara and Fabian Mundt (2017).
  factoextra: Extract and Visualize the Results of
  Multivariate Data Analyses. R package version 1.0.5.
  https://CRAN.R-project.org/package=factoextra
  
  David B. Dahl, David Scott, Charles Roosen, Arni
  Magnusson and Jonathan Swinton (2019). xtable: Export
  Tables to LaTeX or HTML. R package version 1.8-4.
  https://CRAN.R-project.org/package=xtable
  
  
  Malika Charrad, Nadia Ghazzali, Veronique Boiteau, Azam
  Niknafs (2014). NbClust: An R Package for Determining
  the Relevant Number of Clusters in a Data Set. Journal
  of Statistical Software, 61(6), 1-36. URL
  http://www.jstatsoft.org/v61/i06/.
  
  Hadley Wickham (2011). The Split-Apply-Combine Strategy
  for Data Analysis. Journal of Statistical Software,
  40(1), 1-29. URL http://www.jstatsoft.org/v40/i01/.
  
  Barret Schloerke, Jason Crowley, Di Cook, Francois
  Briatte, Moritz Marbach, Edwin Thoen, Amos Elberg and
  Joseph Larmarange (2018). GGally: Extension to
  'ggplot2'. R package version 1.4.0.
  https://CRAN.R-project.org/package=GGally
  
  Alboukadel Kassambara (2019). ggcorrplot: Visualization
  of a Correlation Matrix using 'ggplot2'. R package
  version 0.1.3.
  https://CRAN.R-project.org/package=ggcorrplot
 
  Simon Thuleau and Francois Husson (2019).
  FactoInvestigate: Automatic Description of Factorial
  Analysis. R package version 1.4.
  https://CRAN.R-project.org/package=FactoInvestigate


  
```{r, echo=FALSE, results=FALSE}
Packages <- c("FactoMineR", "factoextra", "xtable", "NbClust", "plyr","GGally", "ggcorrplot", "FactoInvestigate") 
lapply(Packages, citation)
```
